\begin{thebibliography}{00}

\bibitem{whopollution}
World Health Organization. \textit{Air pollution}. Available at: \url{https://www.who.int/health-topics/air-pollution#tab=tab_1}

\bibitem{state}
State of global air. \textit{State of Global Air Report 2024}. Available at: \url{https://www.stateofglobalair.org/resources/report/state-global-air-report-2024?utm_source=chatgpt.com}

\bibitem{aqicn}
AQICN API Documentation. World Air Quality Index Project. (2024). \textit{Real-time Air Quality Index API}. Available at: \url{https://aqicn.org/api/}

\bibitem{google}
Google developers. \textit{API de Air Quality}. Available at: \url{https://developers.google.com/maps/documentation/air-quality/overview?utm_source=chatgpt.com&hl=es-419}

\bibitem{iqair}
IQAir. \textit{Air quality API}. Available at: \url{https://www.iqair.com/us/air-quality-monitors/api?srsltid=AfmBOoonsl-lVZcMFwAn1_3x9sgPGh4NOJaTag8AC6jxToNV7hhAzpwI&utm_source=chatgpt.com}

\bibitem{wiley}
Wiley Advanced. \textit{Advances in Air Quality Monitoring: A Comprehensive Review of Algorithms for Imaging and Sensing Technologies}. Available at: \url{https://advanced.onlinelibrary.wiley.com/doi/full/10.1002/adsr.202300207?utm_source=chatgpt.com}

\bibitem{mdpi}
MDPI. \textit{A Distributed Real-Time Monitoring Scheme for Air Pressure Stream Data Based on Kafka}. Available at: \url{https://www.mdpi.com/2076-3417/14/12/4967?utm_source=chatgpt.com}

\bibitem{medium}
Medium. \textit{Building an IoT Monitoring System with Spark Structured Streaming, Kafka and Scala}. Available at: \url{https://medium.com/%40jr.vera.ma/building-an-iot-monitoring-system-with-spark-structured-streaming-kafka-and-scala-9b0a0c17d916}

\bibitem{esr}
Journal of Electrical Systems. \textit{Real-Time IoT Sensor Data Streaming and Processing with Apache Flink: A Scalable Solution for Smart Monitoring}. Available at: \url{https://journal.esrgroups.org/jes/article/download/8042/5464/14695}

\bibitem{timescale}
Timescale Inc. \textit{Tables and Hypertables}. Available at: \url{https://docs.timescale.com/getting-started/latest/tables-hypertables/}

\bibitem{postgsmv}
PostgreSQL Global Development Group. \textit{REFRESH MATERIALIZED VIEW}. Available at: \url{https://www.postgresql.org/docs/current/sql-refreshmaterializedview.html}

\bibitem{minio}
MinIO Inc. \textit{MinIO Object Storage for Linux}. Available at: \url{https://min.io/docs/minio/linux/index.html}

\bibitem{grafana}
Grafana Labs. \textit{Dashboards Overview}. Available at: \url{https://grafana.com/docs/grafana/latest/fundamentals/dashboards-overview/}

\bibitem{whoaq}
World Health Organization. \textit{Global Air Quality Guidelines}. (2021). Available at: \url{https://www.who.int/publications/i/item/9789240034228}

\bibitem{epaaqi}
U.S. Environmental Protection Agency. \textit{AQI Breakpoints}. Available at: \url{https://aqs.epa.gov/aqsweb/documents/codetables/aqi_breakpoints.html}

\bibitem{jmeter}
Apache Software Foundation. \textit{Apache JMeter User's Manual}. Available at: \url{https://jmeter.apache.org/usermanual/}

\bibitem{googleair}
Google LLC. \textit{Air Quality API Overview}. Available at: \url{https://developers.google.com/maps/documentation/air-quality/overview}

\bibitem{iqair}
IQAir. \textit{AirVisual API Documentation}. Available at: \url{https://www.iqair.com/air-pollution-data-api}

\bibitem{aqicnhist}
World Air Quality Index Project. \textit{Historical CSV Archive}. Available: \url{https://aqicn.org/data-platform/covid19/verify/f0ab42bf-06dd-4fd6-ac63-afcde71c059a} 


\end{thebibliography}